Collaborative Visual Analysis on the Web with RCloud

### 2015-06-25 discussion.tex
fix table columns
fixed spelling errors

\begin{table}
  \centering
  \begin{tabular}{l|cccccc}
    & V./F. & C. & D. & M.L. & R. & A. \\
    \hline
RCloud     &       x            &       x       &     x      &      x        &           x        &           x         \\
RStudio    &                    &               &            &               &           x        &           x         \\
JSFiddle   &       x            &       x       &            &               &                    &                     \\
bl.ocks    &       x            &       x       &            &               &           x        &                     \\
shiny      &                    &               &     x      &               &           x        &           x         \\
Jupyter    &                    &               &            &      x        &           x        &           x         \\
Tableau    &                    &      x        &     x      &               &           x        &                     \\
ManyEyes   &                    &      x        &     x      &               &                    &                     
  \end{tabular}
  \caption{Comparison of system features. Columns: Versioning/Forking, Collaboration, Deployment, Multi-language support, Integrated Reports, and Integrated Analysis.}
\end{table}
% SN says Tableau: versioning=no  collaboration=yes  deployment=yes multilanguage=no  reports=yes  analysis=no

@@

\paragraph*{Shared artifacts and artifact histories} are a central feature
of RCloud, through shared workbooks. We observed that hackers readily
share experiments, demonstrate techniques, publish results to peers and
managers, and transfer algorithms to other groups. Artifact histories
can be accessed through the notebook tree or through GitHub's web interface.
But a probem is that it is easy to create a large number of artifacts
and the system does not help enough to organize and navigate them.

@@

Versioning of different software pieces must be managed in several
levels: in data analysis and visualization scripts (that is, the
notebooks themselves); in the R environment such as the installed
libraries and packages; and in the external environment such as the
operating system, libraries, and protocols spoken by remote
services. At one end of the scale, we have full control over the
versions via git, along with conventions that fix stable notebook
versions.  The R environment itself is under the control of its
package manager that has rules to ensure reasonable consistency.  It
is at least possible for RCloud to access information about package
versions and configurations, but support for checking compatibility
and maintaining multiple versions in the same environment is not
strong.  At the other end of the scale, there is not much reason to
expect version control for the external environment, and consider what
would happend if a web service that is used by a notebook changes the
format of its output: notebooks now stop working, but nothing in
RCloud has changed.

>>>>

\begin{table}
  \centering
  \begin{tabular}{l|cccccc}
    & V/F & C & D & ML & R & A \\
    \hline
RCloud     &       x            &       x       &     x      &      x        &           x        &           x         \\
RStudio    &                    &               &            &               &           x        &           x         \\
JSFiddle   &       x            &       x       &            &               &                    &                     \\
bl.ocks    &       x            &       x       &            &               &           x        &                     \\
shiny      &                    &               &     x      &               &           x        &           x         \\
Jupyter    &                    &               &            &      x        &           x        &           x         \\
Tableau    &                    &      x        &     x      &               &           x        &                     \\
ManyEyes   &                    &      x        &     x      &               &                    &                     
  \end{tabular}
  \caption{Comparison of system features. Columns: Versioning/Forking, Collaboration, Deployment, Multi-language support, Integrated Reports, and Integrated Analysis.}
\end{table}
% SN says Tableau: versioning=no  collaboration=yes  deployment=yes multilanguage=no  reports=yes  analysis=no

@@

\paragraph*{Shared artifacts and artifact histories} are a central feature
of RCloud, through shared workbooks. We observed that hackers readily
share experiments, demonstrate techniques, publish results to peers and
managers, and transfer algorithms to other groups. Artifact histories
can be accessed through the notebook tree or through GitHub's web interface.
But a problem is that it is easy to create a large number of artifacts
and the system does not help enough to organize and navigate them.

@@

Versioning of different software pieces must be managed in several
levels: in data analysis and visualization scripts (that is, the
notebooks themselves); in the R environment such as the installed
libraries and packages; and in the external environment such as the
operating system, libraries, and protocols spoken by remote
services. At one end of the scale, we have full control over the
versions via git, along with conventions that fix stable notebook
versions.  The R environment itself is under the control of its
package manager that has rules to ensure reasonable consistency.  It
is at least possible for RCloud to access information about package
versions and configurations, but support for checking compatibility
and maintaining multiple versions in the same environment is not
strong. At the other end of the scale, there is not much reason to
expect version control for the external environment, and consider what
would happened if a web service that is used by a notebook changes the
format of its output: notebooks now stop working, but nothing in
RCloud has changed.

### 2015-06-25 introduction.tex
fixed spelling errors

\begin{enumerate}
\item Discoverability. A chronic complaint of data analysis teams is
inefficiency due to a lack of transparency in accessing the work of
other team memnbers.  The processes and technologies supporting
data analysis and visual analytics today are fragmented.
Knowledge is shared imperfectly through informal conversations, emails,
meetings, phone calls, source code, wikis and project documents.
This causes work to be repeated unnecessarily
(``How can I connect to database X?'', ``How do you get clean data
from column Y from feed Z?'') due to lack of communication about
previously solved problems. 
Our goal is support and
promote transparency of work between team members.

@@

\item Technology transfer.
In most organizations, development of analyses and visualizations 
s done by \emph{hackers} and \emph{scripters} (adopting Kandel et al.'s
terminology~\cite{Kandel:2012:EDA}). \emph{Application users} 
gain the benefits of the tools developed by hackers and scripters.
Results of experiments are often shared first by copying screen
shots or static output.  By the time decisions are made based on
the results of that screenshot, data and processes may have changed
so much that the decision can be wrong.
\stephen{Transitional sentence here.}
Generally, the connection between these two worlds is made by IT staff,
who port or rewrite code so it can run as a stable production service.
In an environment where business needs can change rapidly, this
process does not scale. It is also difficult to trace problems
from production back to EDA.
Our goal is to merge the worlds of EDA and production to close this gap.

@@

Fortunately, there is abundant evidance that technology can
solve many of these problems.
Gutierrez collected interviews with data scientists~\cite{Gutierrez:2014:DSA},
which include the following:

@@

In consideration of the opportunity to improve this situation,
we created a software environment that supports the end-to-end
visual analytics process for individuals and teams and
their work within larger organizations.

In this paper, we contribute the
design of RCloud, with an interview study based on the
course of its development and deployment at AT\&T Labs for 
about three years.

This environment knits together some familiar tools, and 
provides new features to find data and code; create experimental workbooks;
run experiments; annotate and deploy experiments as end-user websites or
as reusable, callable services; and to share, search and recommend these artifacts. The artifacts are stored in a version
control system that provides a common workspace, as well as needed control
and isolation between stable and experimental versions of code and other
resources.

>>>>

\begin{enumerate}
\item Discoverability. A chronic complaint of data analysis teams is
inefficiency due to a lack of transparency in accessing the work of
other team members. The processes and technologies supporting
data analysis and visual analytics today are fragmented.
Knowledge is shared imperfectly through informal conversations, emails,
meetings, phone calls, source code, wikis and project documents.
This causes work to be repeated unnecessarily
(``How can I connect to database X?'', ``How do you get clean data
from column Y from feed Z?'') due to lack of communication about
previously solved problems. 
Our goal is support and
promote transparency of work between team members.

@@

\item Technology transfer.
In most organizations, development of analyses and visualizations 
s done by \emph{hackers} and \emph{scripters} (adopting Kandel et al.'s
terminology~\cite{Kandel:2012:EDA}). \emph{Application users} 
benefit from tools developed by hackers and scripters.
Results of experiments are often shared first by copying screen
shots or static output. By the time decisions are made based on
the results of those screenshots, data and processes may have changed
so much that the decision can be wrong.
\stephen{Transitional sentence here.}
Generally, the connection between these two worlds is made by IT staff,
who port or rewrite code so it can run as a stable production service.
In an environment where business needs can change rapidly, this
process does not scale. It is also difficult to trace problems
from production back to EDA.
Our goal is to merge the worlds of EDA and production to close this gap.

@@

Fortunately, there is abundant evidence that technology can
solve many of these problems.
Gutierrez collected interviews with data scientists~\cite{Gutierrez:2014:DSA},
which include the following:

@@

In view of the opportunity to improve this situation,
we created a software environment that supports the end-to-end
visual analytics process for individuals and teams and
their work within larger organizations.

In this paper, we contribute the design of RCloud, and an
interview study based on its development and deployment
at AT\&T Labs for about three years.

RCloud knits together some familiar tools, and 
provides new features to find data and code; create experimental workbooks;
run experiments; annotate and deploy experiments as end-user websites or
as reusable, callable services; and to share, search and recommend these artifacts. The artifacts are stored in a version
control system that provides a common workspace, as well as needed control
and isolation between stable and experimental versions of code and other
resources.
